---
title: "The 'qspray' package"
output: github_document
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/stla/qspray/workflows/R-CMD-check/badge.svg)](https://github.com/stla/qspray/actions)
<!-- badges: end -->

*R package for multivariate polynomials with rational coefficients.*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
```

```{r}
library(qspray)
```

Define a polynomial:

```{r}
x <- lone(1); y <- lone(2); z <- lone(3)
pol <- 4*x^2 + "1/2"*y - 5*x*y*z
pol
```

Some arithmetic on this polynomial:

```{r}
-pol
2 * pol
# TODO: pol / 2
"5/3" * pol
pol + 5
pol - "2/5"
pol^2
```

Two polynomials can be added and multiplied:

```{r}
pol1 <- pol
pol2 <- pol
pol1 + pol2
pol1 - pol2
pol1 * pol2
```

Use `evalQspray` to evaluate a polynomial for some values of the variables:

```{r}
evalQspray(pol, c("1", "2", "3/2"))
```

The package also provides a function which returns the exact value of the 
integral of a polynomial with rational coefficients over a simplex with 
rational vertices:

```{r}
# variables
x <- lone(1)
y <- lone(2)
z <- lone(3)
# polynomial
P <- x^4 + y + 2*x*y^2 - 3*z
# simplex (tetrahedron) vertices
v1 <- c(1, 1, 1)
v2 <- c(2, 2, 3)
v3 <- c(3, 4, 5)
v4 <- c(3, 2, 1)
# simplex
S <- rbind(v1, v2, v3, v4)
# integral
integratePolynomialOnSimplex(P, S)
```